---
title: "Chapter 10, Vignette 1: Quick Start: Running DBMH-ORH analysis"
author: "Dev P. Chakraborty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start: Running DBMH-ORH analysis}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(RJafroc)
```

## Introduction
This vignette is intended for those seeking a quick transition from Windows *JAFROC* to *RJafroc*.
Described first is the structure of an `RJafroc` dataset followed by how to read
a *JAFROC*  format file to create an `RJafroc` dataset.

## An ROC dataset
`{dataset3}` corresponding to the Franken ROC data is predefined.
The following code shows the structure of this dataset.

```{r}
str(dataset03)
```             

It is a list with 8 members. The false positive ratings are contained in `{NL}`, an array
with dimensions `[1:2,1:4,1:100,1]`. The first index corresponds to treatments, and since the 
dataset has 2 treatments, the corresponding dimension is 2. The second index corresponds to 
readers, and since the dataset has 4 readers, the corresponding dimension is 4. The third index 
corresponds to the total number of cases. Since the dataset has 100 cases, the corresponding 
dimension is 100. But, as you can see from the code below, the entries in this array for cases 34
through 100 are `-Inf`. 

```{r}
dataset03$NL[1,1,34:100,1]
``` 

This is because in the ROC paradigm false positive are not possible on diseased cases. So the actual
FP ratings are contained in the first 33 elements of the array. How did I know that there are 34
non-diseased cases? This can be understood in several ways.

* `LL` is an array with dimensions `[1:2,1:4,1:67,1]`. This implies 67 diseased cases, and by 
    subtraction from 100, there must be 34 non-diseased cases.
* The list members `lesionNum`, `lesionID` and `lesionWeight` are all arrays with dimensions
    `[1:67,1]` containing ones. Again, these imply 67 diseased cases.
* The fields `lesionNum`, `lesionID` and `lesionWeight`, while not needed for ROC data, 
    are needed, as we shall see later, for the FROC paradigm.

The `dataType` list member is the character string `"ROC"`,
    characterizing the ROC dataset. Alternatives are `"FROC"` and `"LROC"`.
    
```{r}
dataset03$dataType
``` 

The `modalityID` list member is a character string with two entries, "0" and "1",
    corresponding to the two treatments (i.e., modalities). These can be longer
    strings, if you  please, that label the two treatments.
    
```{r}
dataset03$modalityID
``` 

The `readerID` list member is a character string with four entries, "0",  "1", "2" and "3"
    corresponding to the four readers. These can be longer
    strings that label the four readers.
```{r}
dataset03$readerID
``` 


Here are the actual ratings for cases 1:34.


```{r}
dataset03$NL[1,1,1:34,1]
``` 

This says that for treatment 1 and reader 1, (non-diseased) case 1 was rated 3, case 3 was rated 2,
case 8 was rated 4, etc. The `-Inf` corresponds to the cases rated 1, which is is equivalent to a 1-rating.
The reason for this is that the ratings are ordered labels. As far as the ordering is concerned, nothing is
changed be replacing `-Inf` with 1 and vice-versa.

As another example, for treatment 2 and reader 3, 

```{r}
dataset03$NL[2,3,1:34,1]
``` 

As you can see, there are no cases that are explicitly rated 1, so changing the `-Inf` to 1 does not change 
the ordering of the ratings.

## Creating a dataset from a JAFROC format file

There is a file `includedRocData.xlsx` that is part of the package intallation. Since it is a system
file one must get its name as follows.

```{r}
fileName <- "includedRocData.xlsx"
sysFileName <- system.file(paste0("extdata/",fileName), package = "RJafroc", mustWork = TRUE)
``` 

Next, one uses DfReadDataFile() as follows, assuming it is a JAFROC format file.

```{r}
ds <- DfReadDataFile(sysFileName)
``` 

Now ds is the desired dataset.

```{r}
str(ds)
``` 

Analysis is illustrated for `dataset03`, but one could have used the newly created 
dataset `ds`.

## Analyzing the ROC dataset

This illustrates the `StSignificanceTesting()` function. The significance testing method is specified as `"DBMH"`
and the figure of merit `FOM` is specified as "Wilcoxon".

```{r}
ret <- StSignificanceTesting(dataset03, method = "DBMH", FOM = "Wilcoxon")
print(ret)
```             

## Explanation of the output
The function returns a long unwieldy list. Let us consider them one by one.
The function `UtilOutputReport()` can generate an Excel file report, making it much
easier to visualize the results. This is described in another vignette.

### FOMs
* `fomArray` contains the `[1:2,1:4]` FOM values.

```{r}
ret$fomArray
```             


This shows the 2 x 4 array of FOM values. 

### Pseudovalue ANOVA table

* `anovaY`, where the Y denotes that these are pseudovalue based, 
is the ANOVA table.


```{r}
ret$anovaY
```             

### Pseudovalue ANOVA table, each treatment

* `anovaYi` is the ANOVA table for individual treatments.


```{r}
ret$anovaYi
```             

The 0 and 1 headers come from the treatment names.

### Pseudovalue Variance Components

* `varComp` is the variance components (needed for sample
size estimation).


```{r}
ret$varComp
```             

### Random-reader random-case (RRRC) analysis

* `fRRRC` is the F-statistic for testing the NH that the
treatments have identical FOMs. RRRC means random-reader random-case generalization.


```{r}
ret$fRRRC
```             

#### F-statistic and p-value for RRRC analysis
* `ddffRRRC` is the denominator degrees of freedom of the 
F-statistic.


```{r}
ret$ddffRRRC
```             

* `pRRRC` is the p-value of the test.


```{r}
ret$pRRRC
```             

#### Confidence Intervals for RRRC analysis

* `ciDiffTrtRRRC` is the 95% confidence interval of
reader-averaged differences between treatments.


```{r}
ret$ciDiffTrtRRRC
```             

* `ciAvgRdrEachTrtRRRC` is the 95% confidence interval of
reader-averaged FOMs for each treatments.


```{r}
ret$ciAvgRdrEachTrtRRRC
```             

### Fixed-reader random-case (FRRC) analysis

#### F-statistic and p-value for RRRC analysis

* `fFRRC` is the F-statistic for fixed-reader
random-case analysis.


```{r}
ret$fFRRC
```             

* `ndf` is the numerator degrees of freedom of the F-statistic,
always one less than the number of treatments.


```{r}
ret$ndf
```             

* `ddfFRRC` is the denominator degreesof freedom of the F-statistic,
for fixed-reader random-case analysis.


```{r}
ret$ddfFRRC
```             

* `pFRRC` is the p-value for fixed-reader random-case analysis.


```{r}
ret$pFRRC
```             

#### Confidence Intervals for FRRC analysis

* `ciDiffTrtFRRC` is the 95% CI of reader-average
differences between treatments for fixed-reader random-case analysis


```{r}
ret$ciDiffTrtFRRC
```             

* `ciAvgRdrEachTrtRRFC` is the 95% CI of reader-average
FOMs of each treatment for fixed-reader random-case analysis


```{r}
ret$ciAvgRdrEachTrtRRFC
```             

#### ANOVA for FRRC analysis

* `ssAnovaEachRdr` is the sum of squares ANOVA
for each reader


```{r}
ret$ssAnovaEachRdr
```             

* `msAnovaEachRdr` is the mean squares ANOVA
for each reader


```{r}
ret$msAnovaEachRdr
```             

#### Confidence Intervals for FRRC analysis

* `ciDiffTrtFRRC` is the CI for reader-averaged
treatment differences, for fixed-reader random-case analysis


```{r}
ret$ciDiffTrtFRRC
```             

### Random-reader fixed-case (RRFC) analysis

#### F-statistic and p-value for RRFC analysis
* `fRRFC` is the F-statistic for
for random-reader fixed-case analysis


```{r}
ret$fRRFC
```             

* `ddfRRFC` is the ddf for for random-reader fixed-case analysis


```{r}
ret$ddfRRFC
```             

* `pRRFC` is the p-value for for random-reader fixed-case analysis


```{r}
ret$pRRFC
```             

#### Confidence Intervals for RRFC analysis

* `ciDiffTrtRRFC` is the CI for reader-averaged
inter-treatment FOM differences for random-reader fixed-case analysis


```{r}
ret$ciDiffTrtRRFC
```             

* `ciAvgRdrEachTrtRRFC` is the CI for treatment FOMs 
for each reader for random-reader fixed-case analysis


```{r}
ret$ciAvgRdrEachTrtRRFC
```             

## ORH significance testing
Simply change `method = "DBMH"` to `method = "ORH"`.

```{r}
ret <- StSignificanceTesting(dataset03, method = "ORH", FOM = "Wilcoxon")
str(ret)
```
