---
title: "Chapter 10, Vignette 2: Quick Start DBMH-ORH Analysis with Excel Output"
author: "Dev P. Chakraborty, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
vignette: >
  %\VignetteIndexEntry{Quick Start}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(RJafroc)
```

## Introduction
This vignette illustrates significance testing using the DBMH method. But, instead of the unwieldy output in *QuickStartDBMH.html*, it generates an Excel output file containing the following worksheets:

* `Summary`
* `FOMs`
* `RRRC`
* `FRRC`
* `RRFC`
* `ANOVA`


## Generating the Excel output file

This illustrates the `UtilOutputReport()` function. The significance testing `method` is "DBMH", the default, and the figure of merit `FOM` is  "Wilcoxon". Note `ReportFileExt` = "xlsx"` telling
the function to create an Excel output file. The Excel output is created in a temporary file.

```{r}
ret <- UtilOutputReport(get("dataset03"), FOM = "Wilcoxon", method = "DBMH", overWrite = TRUE, ReportFileExt = "xlsx")
```             


## ORH significance testing
Simply change `method = "DBMH"` (the default) to `method = "ORH"`.


```{r}
ret <- UtilOutputReport(get("dataset03"), FOM = "Wilcoxon", method = "ORH", overWrite = TRUE, ReportFileExt = "xlsx")
```             


