---
title: "Chapter 11, Vignette 1: Background on the F distribution"
author: "Dev P. Chakraborty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ROC sample size}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
```

## Illustrating the F-distribution: `ndf` = 2, `ddf` = 10

```{r, fig.align = "center"}
ndf <- 2;ddf <- 10;ncp <- c(0,2,5,10)
fCrit1 <- qf(0.95, ndf,ddf)
cat("critical value of x for rejecting NH is ", fCrit1,"\n")
x <- seq(1, 20, 0.1)
myLabel <- c("A", "B", "C", "D")
myLabelIndx <- 1
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  cat("ndf = ", ndf, ", ddf = ", ddf, ", ncp = ", ncp[i], ", prob > fCrit1 = ", 
      1-pf(fCrit1, ndf, ddf, ncp = ncp[i]), "\n")
  curveData <- data.frame(x = x, pdf = y)
  curvePlot <- ggplot(data = curveData, mapping = aes(x = x, y = pdf)) + 
    geom_line() +
    ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
  print(curvePlot)
}
```             

### Comments

(A) Corresponds to `ncp = 0`, which yields the usual (aka *central*) F-distribution that was used in the previous two chapters (9 and 10). The integral under this distribution is unity (this is true for all plots above). The critical value, `fCrit1` in the code, is that value of `x` such that the probability of exceeding `x` is $\alpha$; in the current example `fCrit1` = `r qf(0.95, ndf,ddf)`. Notice the use of the quantile function `qf()` to determine this value, also that the default value of `ncp`, namely zero, is used; specifically, one does not pass a 4th argument to `qf()`. **The decision rule for rejecting the NH uses the NH distribution of the F-statistic**, i.e., reject the NH if F >= `fCrit1`. As expected, `prob > fCrit1` = `r 1-pf(fCrit1, ndf, ddf, ncp = ncp[1])` because this is how `fCrit1` was defined. 
(B) Corresponds to `ncp = 2`. It is noticeably shifted to the right as compared to A, thereby making it more likely that the observed value of the F-statistic will exceed the critical value determined for the NH distribution. In fact, `prob > fCrit1` = `r 1-pf(fCrit1, ndf, ddf, ncp = ncp[2])`, which is the *statistical power*. 
(C) Corresponds to `ncp = 5`, and now `prob > fCrit1` = `r 1-pf(fCrit1, ndf, ddf, ncp = ncp[3])`. 
(D) Corresponds to `ncp = 10`, and now `prob > fCrit1` is  `r 1-pf(fCrit1, ndf, ddf, ncp = ncp[4])`. 
The effect of the shift is obvious in plots (C) and (D). Considering a vertical line at `x = 4.10`, most of the distribution in plot (D), more than `r 1-pf(fCrit1, ndf, ddf, ncp = ncp[4])`, to be precise, lies to the right of this line, so the NH is more than `r 1-pf(fCrit1, ndf, ddf, ncp = ncp[4])` likely to be rejected. Of course, what causes the shift is the increased value of the non-centrality parameter. The larger that non-centrality parameter, the greater the shift to the right.

## Illustrating the F-distribution: `ndf` = 2, `ddf` = 100

```{r, fig.align = "center"}
ndf <- 2;ddf <- 100
fCrit2 <- qf(0.95, ndf,ddf)
cat("critical value of x for rejecting NH is ", fCrit2,"\n")
x <- seq(1, 20, 0.1)
myLabel <- c("E", "F", "G", "H")
myLabelIndx <- 1
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  cat("ndf = ", ndf, ", ddf = ", ddf, ", ncp = ", ncp[i], ", prob > fCrit2 = ", 
      1-pf(fCrit2, ndf, ddf, ncp = ncp[i]), "\n")
  curveData <- data.frame(x = x, pdf = y)
  curvePlot <- ggplot(data = curveData, mapping = aes(x = x, y = pdf)) + 
    geom_line() +
    ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
  print(curvePlot)
}
```             

### Comments

(E) Corresponds to `ncp = 0`. The critical value is `fCrit2` = `r qf(0.95, ndf,ddf)`. Notice the decrease compared to the previous value, i.e., `r qf(0.95, ndf,10)`, for ddf = 10. 
(F) Corresponds to `ncp = 2`, and `prob > fCrit2` = `r 1-pf(fCrit2, ndf, ddf, ncp = ncp[2])`, greater than the previous value `r 1-pf(fCrit1, ndf, 10, ncp = ncp[2])`. 
(G) Corresponds to `ncp = 5`, and now `prob > fCrit2` = `r 1-pf(fCrit2, ndf, ddf, ncp = ncp[3])`. 
(H) Corresponds to `ncp = 10`, and now `prob > fCrit2` is  `r 1-pf(fCrit2, ndf, ddf, ncp = ncp[4])`. 

## Illustrating the F-distribution: `ndf` = 1, `ddf` = 100

```{r, fig.align = "center"}
ndf <- 1;ddf <- 100
fCrit3 <- qf(0.95, ndf,ddf)
cat("critical value of x for rejecting NH is ", fCrit3,"\n")
x <- seq(1, 20, 0.1)
myLabel <- c("I", "J", "K", "L")
myLabelIndx <- 1
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  cat("ndf = ", ndf, ", ddf = ", ddf, ", ncp = ", ncp[i], ", prob > fCrit3 = ", 
      1-pf(fCrit3, ndf, ddf, ncp = ncp[i]), "\n")
  curveData <- data.frame(x = x, pdf = y)
  curvePlot <- ggplot(data = curveData, mapping = aes(x = x, y = pdf)) + 
    geom_line() +
    ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
  print(curvePlot)
}
```             

### Comments

(I) Corresponds to `ncp = 0`. The critical value is `fCrit3` = `r qf(0.95, ndf,ddf)`. Notice the increase compared to the corresponding value for `ndf = 2`, i.e., `r qf(0.95, 2,100)`.
(J) Corresponds to `ncp = 2`, and `prob > fCrit3` = `r 1-pf(fCrit3, ndf, ddf, ncp = ncp[2])`, larger than the previous value `r 1-pf(fCrit2, 2, 100, ncp = ncp[2])`. 
(K) Corresponds to `ncp = 5`, and now `prob > fCrit3` = `r 1-pf(fCrit3, ndf, ddf, ncp = ncp[3])`, larger than the previous value `r 1-pf(fCrit2, 2, 100, ncp = ncp[3])`. 
(L) Corresponds to `ncp = 10`, and now `prob > fCrit3` is  `r 1-pf(fCrit3, ndf, ddf, ncp = ncp[4])`, larger than the previous value `r 1-pf(fCrit2, 2, 100, ncp = ncp[4])`. 

## Summary
Power increases with increasing `ddf` and `ncp`; the effect of increasing `ncp` is quite dramatic. Decreasing `ndf` also increases power. 