---
title: "Chapter 11, Vignette 1: Background on the F distribution"
author: "Dev P. Chakraborty, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: myRefs.bib
vignette: >
  %\VignetteIndexEntry{ROC sample size background}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(ggplot2)
```

## Introduction
Since it plays an important role in sample size estimation, it is helpful to examine the behavior of the F-distribution. In the following `ndf` = numerator degrees of freedom, `ddf` = denominator degrees of freedom and `ncp` = non-centrality parameter (i.e., the $\Delta$ appearing in Eqn. (11.6) of [@RN2680]). 

The use of three `R` functions is demonstrated. 

* `qf(p,ndf,ddf)` is the *quantile* function of the F-distribution for specified values of `p`, `ndf` and `ddf`, i.e., the value `x` such that fraction `p` of the area under the F-distribution lies to the right of `x`. Since `ncp` is not included as a parameter, the default value, i.e., zero, is used. This is called the *central* F-distribution. 

* `df(x,ndf,ddf,ncp)` is the probability density function (*pdf*) of the F-distribution, as a function of `x`, for specified values of `ndf`, `ddf` and `ncp`. 

* `pf(x,ndf,ddf,ncp)` is the probability (or cumulative) distribution function of the F-distribution for specified values of `ndf`, `ddf` and `ncp`.


## Effect of `ncp` for `ndf` = 2 and `ddf` = 10
* Four values of `ncp` are considered `r ncp <- c(0,2,5,10)` (`r ncp`) for `ddf` = 10. 
* `fCrit` is the critical value of the F distribution, i.e., that value such that fraction $\alpha$ of the area is to the right of the critical value, i.e., `fCrit` is identical in statistical notation to  ${{F}_{1-\alpha ,ndf,ddf}}$.

```{r, fig.show='hold'}
ndf <- 2;ddf <- 10;ncp <- c(0,2,5,10)
alpha <- 0.05
fCrit <- qf(1-alpha, ndf,ddf)
cat("critical value of x for rejecting NH is ", fCrit,"\n")
x <- seq(1, 20, 0.1)
myLabel <- c("A", "B", "C", "D")
myLabelIndx <- 1
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  cat("ndf = ", ndf, ", ddf = ", ddf, ", ncp = ", ncp[i], ", prob > fCrit = ", 
      1-pf(fCrit, ndf, ddf, ncp = ncp[i]), "\n")
}  
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  curveData <- data.frame(x = x, pdf = y)
  curvePlot <- ggplot(data = curveData, mapping = aes(x = x, y = pdf)) + 
    geom_line() +
    ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
  print(curvePlot)
}
fCrit_2_10 <- fCrit # convention fCrit_ndf_ddf
```             

## Comments

### Fig. (A) 
* This corresponds to `ncp = 0`, i.e., the *central* F-distribution. 
* The integral under this distribution is unity (this is also true for all plots in this vignette). 
* The critical value, `fCrit` in the above code block, is the value of `x` such that the probability of exceeding `x` is $\alpha$. The corresponding parameter `alpha` is defined above as `r alpha`.
* In the current example `fCrit` = `r qf(1-alpha, ndf,ddf)`. Notice the use of the quantile function `qf()` to determine this value, and the default value of `ncp`, namely zero, is used; specifically, one does not pass a 4th argument to `qf()`. 
* **The decision rule for rejecting the NH uses the NH distribution of the F-statistic**, i.e., reject the NH if F >= `fCrit`. As expected, `prob > fCrit` = `r 1-pf(fCrit, ndf, ddf, ncp = ncp[1])` because this is how `fCrit` was defined.   

### Fig. (B) 
* This corresponds to `ncp = 2`, `ndf` = 2 and `ddf` = 10. 
* The distribution is slightly shifted to the right as compared to A, thereby making it more likely that the observed value of the F-statistic will exceed the critical value determined for the NH distribution. 
* In fact, `prob > fCrit` = `r 1-pf(fCrit, ndf, ddf, ncp = ncp[2])`, i.e., the *statistical power* (compare this to Fig. A where `prob > fCrit`  was 0.05).  

### Fig. (C) 
* This corresponds to `ncp = 5`, `ndf` = 2 and `ddf` = 10.
* Now `prob > fCrit` = `r 1-pf(fCrit, ndf, ddf, ncp = ncp[3])`. 
* Power has increased compared to Fig. B.

### Fig. (D) 
* This corresponds to `ncp = 10`, `ndf` = 2 and `ddf` = 10.
* Now `prob > fCrit` is  `r 1-pf(fCrit, ndf, ddf, ncp = ncp[4])`. 
* Power has increased compared to Fig. C.
* The effect of the shift is most obvious in plots (C) and (D). 
* Considering a vertical line at `x ` = `r fCrit_2_10`, fraction `r 1-pf(fCrit, ndf, ddf, ncp = ncp[4])` of the probability distribution in plot (D) lies to the right of this line
* Therefore the NH is likely to be rejected with probability `r 1-pf(fCrit, ndf, ddf, ncp = ncp[4])`. 

### Summary 
The larger that non-centrality parameter, the greater the shift to the right of the F-distribution, and the greater the statistical power.

## Effect of `ncp` for `ndf` = 2 and `ddf` = 100

```{r, fig.show='hold'}
ndf <- 2;ddf <- 100
fCrit <- qf(1-alpha, ndf,ddf)
cat("critical value of x for rejecting NH is ", fCrit,"\n")
x <- seq(1, 20, 0.1)
myLabel <- c("E", "F", "G", "H")
myLabelIndx <- 1
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  cat("ndf = ", ndf, ", ddf = ", ddf, ", ncp = ", ncp[i], ", prob > fCrit = ", 
      1-pf(fCrit, ndf, ddf, ncp = ncp[i]), "\n")
}  
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  curveData <- data.frame(x = x, pdf = y)
  curvePlot <- ggplot(data = curveData, mapping = aes(x = x, y = pdf)) + 
    geom_line() +
    ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
  print(curvePlot)
}
fCrit_2_100 <- fCrit # convention fCrit_ndf_ddf
```             

## Comments

* All comparisons in this sections are at the same values of `ncp` defined above. 
* And between `ddf` = 100 and `ddf` = 10. 

### Fig. (E) 
* This corresponds to `ncp` = 0, `ndf` = 2 and `ddf` = 100. 
* The critical value is `fCrit_2_100` = `r qf(1-alpha, ndf,ddf)`. Notice the decrease compared to the previous value for `ncp` = 0, i.e., `r qf(1-alpha, ndf, 10)`, for `ddf` = 10. 
* One expects that increasing `ddf` will make it more likely that the NH will be rejected, and this is confirmed below.
* All else equal, statistical power increases with increasing `ddf`.

### Fig. (F) 
* This corresponds to `ncp` = 2, `ndf` = 2 and `ddf` = 100. 
* The probability of exceeding the critical value is  `prob > fCrit_2_100` = `r 1-pf(fCrit_2_100, ndf, ddf, ncp = ncp[2])`, greater than the previous value, i.e., `r 1-pf(fCrit_2_10, ndf, 10, ncp = ncp[2])` for `ddf` = 10.

### Fig. (G) 
* This corresponds to `ncp = 5`, `ndf` = 2 and `ddf` = 100.
* The probability of exceeding the critical value is `prob > fCrit_2_100` = `r 1-pf(fCrit_2_100, ndf, ddf, ncp = ncp[3])`.
* This is greater than the previous value, i.e., `r 1-pf(fCrit_2_10, ndf, 10, ncp = ncp[3])` for `ddf` = 10. 

### Fig. (H) 
* This corresponds to `ncp = 10`, `ndf` = 2 and `ddf` = 100. 
* The probability of exceeding the critical value is `prob > fCrit_2_100` is  `r 1-pf(fCrit_2_100, ndf, ddf, ncp = ncp[4])`. 
* This is greater than the previous value, i.e., `r 1-pf(fCrit_2_10, ndf, 10, ncp = ncp[4])` for `ddf` = 10. 

## Effect of `ncp` for `ndf` = 1, `ddf` = 100

```{r, fig.show='hold'}
ndf <- 1;ddf <- 100
fCrit <- qf(1-alpha, ndf,ddf)
cat("critical value of x for rejecting NH is ", fCrit,"\n")
x <- seq(1, 20, 0.1)
myLabel <- c("I", "J", "K", "L")
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  cat("ndf = ", ndf, ", ddf = ", ddf, ", ncp = ", ncp[i], ", prob > fCrit = ", 
      1-pf(fCrit, ndf, ddf, ncp = ncp[i]), "\n")
}  
myLabelIndx <- 1
for (i in 1:length(ncp))
{
  y <- df(x,ndf,ddf,ncp=ncp[i])
  curveData <- data.frame(x = x, pdf = y)
  curvePlot <- ggplot(data = curveData, mapping = aes(x = x, y = pdf)) + 
    geom_line() +
    ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
  print(curvePlot)
}
fCrit_1_100 <- fCrit # convention fCrit_ndf_ddf
```             

## Comments

* All comparisons in this sections are at the same values of `ncp` defined above and at `ddf` = 100.
* And between `ndf` = 1 and `ndf` = 2. 

### Fig. (I) 
* This corresponds to `ncp` = 0, `ndf` = 1 and `ddf` = 100. 
* The critical value is `fCrit_1_100` = `r qf(1-alpha, ndf,ddf)`. 
* Notice the increase in the critical value as compared to the corresponding value for `ndf = 2`, i.e., `r qf(1-alpha, 2,100)`.
* One might expect power to decrease, **but see below**.

### Fig. (J) 
* This corresponds to `ncp` = 2, `ndf` = 1 and `ddf` = 100.
* Now `prob > fCrit_1_100` = `r 1-pf(fCrit_1_100, ndf, ddf, ncp = ncp[2])`, larger than the previous value `r 1-pf(fCrit_2_100, 2, 100, ncp = ncp[2])`.
* The power has actually increased.

### Fig. (K) 
* This corresponds to `ncp` = 5, `ndf` = 1 and `ddf` = 100`', 
* Now `prob > fCrit_1_100` = `r 1-pf(fCrit_1_100, ndf, ddf, ncp = ncp[3])`, larger than the previous value `r 1-pf(fCrit_2_100, 2, 100, ncp = ncp[3])`. 
* Again, the power has actually increased.

### Fig. (L) 
* This corresponds to `ncp` = 10, `ndf` = 1 and `ddf` = 100
* Now `prob > fCrit_1_100` is  `r 1-pf(fCrit_1_100, ndf, ddf, ncp = ncp[4])`, larger than the previous value `r 1-pf(fCrit_2_100, 2, 100, ncp = ncp[4])`. 
* The power has actually increased.

## Summary
* Power increases with increasing `ddf` and `ncp`. 
* The effect of increasing `ncp` is quite dramatic. 
* Decreasing `ndf` also **increases** power. At first glance this may seem counterintuitive, as `fCrit` has gone up, but can be explained by the differing shapes of the two distributions: the pdf is broader for `ndf` = 1 as compared to `ndf` = 2 (compare plot L to H). 

## References
