---
title: "Chapter 11, Vignette 4: RJafroc functions for ROC sample size estimation using ORH method"
author: "Dev P. Chakraborty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{ROC sample size}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(RJafroc)
```

## Introduction
This illustrates the `RJafroc` implementation of sample-size estimation using `"ORH"` method for significance testing.

## Illustration of `SsPowerGivenJK()` using `method = "ORH"`
```{r}
power <- SsPowerGivenJK(dataset02, 6, 251, method = "ORH", option = "RRRC")
````

The returned value is a list containing the expected power, the non-centrality parameter, the denominator degrees of freedom and the F-statistic. 

```{r}
str(power)
````

Expected power is `r power$powerRRRC`.


## Illustration of `SsPowerTable()` using `method = "ORH"`
```{r}
powTab <- SsPowerTable(dataset02, method = "ORH", option = "RRRC")
````

Now show the power table `powTab`.

```{r}
powTab
````

## Illustration of `SsSampleSizeKGivenJ()` using `method = "ORH"`
```{r}
ncases <- SsSampleSizeKGivenJ(dataset02, J = 10, method = "ORH", option = "RRRC")
````

`ncases` is a list containing the number of cases `ncases$K` and expected power `ncases$power`. 

```{r}
str(ncases)
````

The required number of cases is `r ncases$K` and expected power is `r ncases$power`.



