---
title: "Chapter 17, Vignette 1: RSM predicted operating characteristics: lesDist and lesWghtDistr"
author: "Dev P. Chakraborty, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
bibliography: myRefs.bib
fig_caption: yes
fig.width: 4
vignette: >
  %\VignetteIndexEntry{RSM predicted operating characteristics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
  library(RJafroc)
```

## Introduction
* The purpose of this vignette is to explain the operating characteristics predicted by the RSM. It relates to Chapter 17 in my book [@RN2680]. 
* This vignette is under development ...
* Also to explain the difference between `dataset` members (`lesionID`, `lesionWeight`) and (`lesDist`,  `lesWghtDistr`), which are RSM model parameters. 

## The distinction between predicted curves and empirical curves
* Operating characteristics predicted by a model have zero sampling variability.  
* Empirical operating characteristics, which apply to datasets, have non-zero sampling variability. 
* If the model is corect, as the numbers of cases in the dataset increases, the empirical operating characteristic asymptotically approaches the predicted curve. 

## The RSM model
* The 3 RSM parameters and two additional parameters characterizing the dataset determine the wAFROC curve.
* The 3 RSM parameters are $\mu$, $\lambda$ and $\nu$. 
* The two dataset parameters are:
    + The distribution of number of lesions per diseased case, `lesDist`.
    + The distribution of lesion weights, `lesWghtDistr`.
* These parameters do not apply to individual cases; rather they refer to a large population (asymptotically infinite in size) of cases. 

```{r}
str(dataset04$lesions$IDs)
str(dataset04$lesions$weights)
```

* Note that the first index of both arrays is the case index for the 100 abnormal cases in this dataset. 
* With finite number of cases the empirical operating characteristic (or for that matter any fitted operating characteristic) will have sampling variability as in the following example.

## The empirical wAFROC

```{r}
p <- PlotEmpiricalOperatingCharacteristics(dataset04, opChType = "wAFROC")
p$Plot
```

* The piecewise linear nature of the plot, with sharp breaks, indicates that this is due to a finite dataset.
* In contrast the following code shows a smooth plot, because it is a model *predicted* plot.

## The predicted wAFROC

```{r}
## Following example is for mu = 2, lambda = 1, nu = 0.6. 20% of the diseased 
## cases have a single lesion, 40% have two lesions, 10% have 3 lesions, 
## and 30% have 4 lesions.  
lesDistr <- c(0.2, 0.4, 0.1, 0.3)

## On cases with one lesion the weights are 1, on cases with 2 lesions the weights
## are 0.4 and 0.6, on cases with three lesions the weights are 0.2, 0.3 and 0.5, and
## on cases with 4 lesions the weights are 0.3, 0.4, 0.2 and 0.1: 
relWeights <- c(0.3,  0.4, 0.2,  0.1)
p <- PlotRsmOperatingCharacteristics(
  mu = 2, 
  lambda = 1, 
  nu = 0.6, 
  OpChType = "wAFROC", 
  lesDistr = lesDistr, 
  relWeights = relWeights, 
  legendPosition = "bottom", nlfRange = c(0, 1), llfRange = c(0, 1))
p$wAFROCPlot
```

## The distribution of number of lesions and weights 


```{r}
lesDistr
relWeights
```

* The `lesDistr` array `r lesDistr` specifies the fraction of diseased cases with the number of lesions corresponding to the column index. To specify a dataset with exactly 3 lesions per diseased case use `lesDist = c(0, 0, 1, 0)`.
* The `relWeights` array `r relWeights` specifies the relative weights. 
* For cases with 1 lesion, the weight is 1. 
* For cases with 2 lesions, the first lesion has weight `r relWeights[1]/sum(relWeights[1:2])` and the second lesion has weight `r relWeights[2]/sum(relWeights[1:2])`, which are in the ratio 0.3 : 0.4 and sum to unity.
* For cases with 3 lesions, the first lesion has weight `r relWeights[1]/sum(relWeights[1:3])`, the second lesion has weight `r relWeights[2]/sum(relWeights[1:3])` and the third lesion has weight `r relWeights[3]/sum(relWeights[1:3])`, which are in the ratio 0.3 : 0.4 : 0.2, and sum to unity. 
* For cases with 4 lesions, the weights are `r relWeights[1]/sum(relWeights[1:4])`, `r relWeights[2]/sum(relWeights[1:4])`, `r relWeights[3]/sum(relWeights[1:4])` and `r relWeights[4]/sum(relWeights[1:4])`, which are  in the ratio 0.3 : 0.4 : 0.2 : 0.1 and sum to unity.


## Other operating characteristics
* By changing `OpChType` one can generate other operating characteristics.
* Note that lesiion weights argument is not needed for ROC curves. It is only needed for `wAFROC` and `wAFROC1` curves.

```{r}
lesDistr <- c(0.2, 0.4, 0.1, 0.3)
p <- PlotRsmOperatingCharacteristics(
  mu = 2, 
  lambda = 1, 
  nu = 0.6, 
  OpChType = "ROC",
  lesDistr = lesDistr,  
  legendPosition = "bottom")
p$ROCPlot
```


## Summary  



## References  
