---
title: "Chapter 19, Vignette 2: FROC sample size estimation using specified ROC effect size"
author: "Dev P. Chakraborty, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
bibliography: myRefs.bib
vignette: >
  %\VignetteIndexEntry{FROC sample size 2}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
library(RJafroc)
library(ggplot2)
```

## Introduction
This example uses the FED dataset as a pilot FROC study and function `SsFrocNhRsmModel()` to construct the NH model (encapsulating some of the code in the previous vignette).

## Constructing the NH model for the dataset
One starts by extracting the first two treatments from `dataset04`, which represent the NH dataset, see previous vignette. Next one constructs the NH model - note that the lesion distribution `lesionPmf` can be specified here independently of that in the pilot dataset. This allows some control over selection of the diseased cases in the pivotal study.

```{r} 
frocNhData <- DfExtractDataset(dataset04, trts = c(1,2))
ret <- SsFrocNhRsmModel(frocNhData, lesionPmf = c(0.7, 0.2, 0.1))
muMed <- ret$muMed
lambdaMed <- ret$lambdaMed
nuMed <- ret$nuMed
lesDistr <- ret$lesDistr
lesWghtDistr <- ret$lesWghtDistr
scaleFactor <- ret$scaleFactor
````


The fitting model is defined by `muMed = r muMed`,  `lambdaMed = `r lambdaMed``  and  `nuMed = `r nuMed`` and `lesionPmf`. The effect size scale factor is `r scaleFactor`.


```{r}
aucRocNH <- PlotRsmOperatingCharacteristics(muMed, lambdaMed, nuMed, 
                                            lesDistr = lesDistr, 
                                            lesWghtDistr = lesWghtDistr, type = "ROC")$aucROC
aucwAfrocNH <- PlotRsmOperatingCharacteristics(muMed, lambdaMed, nuMed, 
                                               lesDistr = lesDistr, 
                                               lesWghtDistr = lesWghtDistr, type = "wAFROC")$aucwAFROC
````

The null hypothesis ROC AUC is `r aucRocNH` and the corresponding NH wAFROC AUC is `r aucwAfrocNH`. 

## Extracting the wAFROC variance components

The next code block applies `StSignificanceTesting()` to `frocNhData`, using `FOM = "wAFROC"` and extracts the variance components.

```{r}
varCompwAFROC  <- StSignificanceTesting(frocNhData, FOM = "wAFROC", method = "DBMH", option = "RRRC")$varComp
````


## wAFROC power for specified ROC effect size, number of readers J and number of cases K

The following example is for ROC effect size = 0.05, 5 readers (`J`) and 100 cases (`K`) in the **pivotal study**. 


```{r}
ROC_ES <- 0.05
effectSizewAFROC <- scaleFactor * ROC_ES
J <- 5;K <- 100

varYTR <- varCompwAFROC$varTR 
varYTC <- varCompwAFROC$varTC
varYEps <- varCompwAFROC$varErr
ret <- SsPowerGivenJKDbmVarComp (J = J, K = K, effectSize = effectSizewAFROC, 
                                 varYTR, varYTC, varYEps, option = "RRRC")
powerwAFROC <- ret$powerRRRC
  
cat("ROC-ES = ", ROC_ES, ", wAFROC-ES = ", ROC_ES * scaleFactor, ", Power-wAFROC = ", powerwAFROC, "\n")
````

## wAFROC number of cases for 80% power for a given number of readers J


```{r}

varYTR <- varCompwAFROC$varTR 
varYTC <- varCompwAFROC$varTC
varYEps <- varCompwAFROC$varErr
ret2 <- SsSampleSizeKGivenJ(dataset = NULL, J = 6, effectSize = effectSizewAFROC, method = "DBMH", 
                      list(varYTR = varYTR, varYTC = varYTC, varYEps = varYEps))

cat("ROC-ES = ", ROC_ES, ", wAFROC-ES = ", ROC_ES * scaleFactor, 
    ", K80RRRC = ", ret2$KRRRC, ", Power-wAFROC = ", ret2$powerRRRC, "\n")
````


## wAFROC Power for a given number of readers J and cases K


```{r}

ret3 <- SsPowerGivenJK(dataset = NULL, J = 6, K = ret2$KRRRC, effectSize = effectSizewAFROC, method = "DBMH", 
                    list(varYTR = varYTR, varYTC = varYTC, varYEps = varYEps))

cat("ROC-ES = ", ROC_ES, ", wAFROC-ES = ", ROC_ES * scaleFactor, 
    ", powerRRRC = ", ret3$powerRRRC, "\n")
````


The estimated power is close to 80% as the number of cases (`ret2$KRRRC = `r ret2$KRRRC``) was chosen deliberately from the previous code block.


## References


