---
title: "Chapter 19, Vignette 2: FROC sample size estimation using specified ROC effect size"
author: "Dev P. Chakraborty, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_document
bibliography: myRefs.bib
vignette: >
  %\VignetteIndexEntry{Chapter 19, Vignette 2: FROC sample size estimation using specified ROC effect size}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
knitr::opts_chunk$set(
collapse = TRUE,
comment = "#>"
)
library(RJafroc)
library(ggplot2)
```

## Introduction
This example uses the FED dataset as a pilot FROC study and function `SsFrocNhRsmModel()` to construct the NH model (encapsulating some of the code in the previous vignette).

## Constructing the NH model for the dataset
One starts by extracting the first two treatments from `dataset04`, which represent the NH dataset, see previous vignette. Next one constructs the NH model - note that the lesion distribution `lesionPmf` can be specified here independently of that in the pilot dataset. This allows some control over selection of the diseased cases in the pivotal study.

```{r} 
lesDistr <- c(0.7, 0.2, 0.1)
frocNhData <- DfExtractDataset(dataset04, trts = c(1,2))
ret <- SsFrocNhRsmModel(frocNhData, lesDistr = lesDistr)
muMed <- ret$muMed
lambdaMed <- ret$lambdaMed
nuMed <- ret$nuMed
scaleFactor <- ret$scaleFactor
````


The fitting model is defined by `muMed = r muMed`,  `lambdaMed = `r lambdaMed``  and  `nuMed = `r nuMed`` and `lesionPmf`. The effect size scale factor is `r scaleFactor`.


```{r}
aucRocNH <- PlotRsmOperatingCharacteristics(muMed, lambdaMed, nuMed, 
                                            lesDistr = lesDistr, OpChType = "ROC")$aucROC
aucwAfrocNH <- PlotRsmOperatingCharacteristics(muMed, lambdaMed, nuMed, 
                                               lesDistr = lesDistr, OpChType = "wAFROC")$aucwAFROC
````

The null hypothesis ROC AUC is `r aucRocNH` and the corresponding NH wAFROC AUC is `r aucwAfrocNH`. 

## Extracting the wAFROC variance components

The next code block applies `StSignificanceTesting()` to `frocNhData`, using `FOM = "wAFROC"` and extracts the variance components.

```{r}
varCompwAFROC  <- StSignificanceTesting(frocNhData, FOM = "wAFROC", method = "DBM", analysisOption = "RRRC")$ANOVA$VarCom
````


## wAFROC power for specified ROC effect size, number of readers J and number of cases K

The following example is for ROC effect size = 0.05, 5 readers (`J`) and 100 cases (`K`) in the **pivotal study**. 


```{r}
ROC_ES <- 0.05
effectSizewAFROC <- scaleFactor * ROC_ES
J <- 5;K <- 100

varYTR <- varCompwAFROC["VarTR","Estimates"] 
varYTC <- varCompwAFROC["VarTC","Estimates"]
varYEps <- varCompwAFROC["VarErr","Estimates"]
ret <- SsPowerGivenJK(dataset = NULL, FOM = "Wilcoxon", J = J, K = K, analysisOption = "RRRC", 
                      effectSize = effectSizewAFROC, method = "DBM", LegacyCode = TRUE, 
                      list(VarTR = varYTR,
                           VarTC = varYTC,
                           VarErr = varYEps))
powerwAFROC <- ret$powerRRRC

cat("ROC-ES = ", ROC_ES, ", wAFROC-ES = ", ROC_ES * scaleFactor, ", Power-wAFROC = ", powerwAFROC, "\n")
````

## wAFROC number of cases for 80% power for a given number of readers J


```{r}
VarTR <- varCompwAFROC["VarTR","Estimates"] 
VarTC <- varCompwAFROC["VarTC","Estimates"]
VarErr <- varCompwAFROC["VarErr","Estimates"]
ret2 <- SsSampleSizeKGivenJ(dataset = NULL, J = 6, effectSize = effectSizewAFROC, method = "DBM", LegacyCode = TRUE,
                            list(VarTR = VarTR, VarTC = VarTC, VarErr = VarErr))

cat("ROC-ES = ", ROC_ES, ", wAFROC-ES = ", ROC_ES * scaleFactor, 
    ", K80RRRC = ", ret2$KRRRC, ", Power-wAFROC = ", ret2$powerRRRC, "\n")
````


## wAFROC Power for a given number of readers J and cases K


```{r}

ret3 <- SsPowerGivenJK(dataset = NULL, J = 6, K = ret2$KRRRC, effectSize = effectSizewAFROC, method = "DBM", LegacyCode = TRUE,
                       list(VarTR = VarTR, VarTC = VarTC, VarErr = VarErr))

cat("ROC-ES = ", ROC_ES, ", wAFROC-ES = ", ROC_ES * scaleFactor, 
    ", powerRRRC = ", ret3$powerRRRC, "\n")
````


The estimated power is close to 80% as the number of cases (`ret2$KRRRC = `r ret2$KRRRC``) was chosen deliberately from the previous code block.


## References


