---
title: "Chapter 20, Vignette 4: Metz Eqn36 numerical check"
author: "Dev P. Chakraborty, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
fig.width: 4
vignette: >
  %\VignetteIndexEntry{Chapter 20, Vignette 4: Metz Eqn36 numerical check}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
  library(RJafroc)
  library(ggplot2)
  library(mvtnorm)
```

## Helper functions
```{r echo=FALSE}
Transform2ab <- function (d_a, c){
  b <- -(c+1)/(c-1)
  a <- (d_a/sqrt(2))*sqrt(1+b^2)
  return( list(
    a = a,
    b = b
  ) )
}

GetLimits <- function (d_a, c)
{
  if (c < 0.0) {
    LL <-  d_a/4/c*sqrt(1+c^2)
    UL <-  10.0
  }
  if (c == 0.0) 
  {
    LL  <-  -10.0
    UL <-  10.0
  }
  if (c > 0.0) 
  {
    LL <-  -10.0
    UL <-  d_a/4/c*sqrt(1+c^2)
  }
  return( list(
    LL = LL,
    UL = UL
  ) )
}

Heaviside <- function (c)
{
  if (c < 0.0) rval <-  0.0 else rval  <-  1.0
  return (rval)
}

FalsePositiveFraction <- function (vc, d_a, c)
{
  
  arg1 <-  -(1-c)*vc-d_a/2*sqrt(1+c^2)
  arg2 <-  -(1-c)*vc+d_a/2/c*sqrt(1+c^2)
  
  rval = pnorm (arg1) + pnorm (arg2) - Heaviside (c)
  
  return (rval)
}

TruePositiveFraction <- function (vc, d_a, c)
{
  arg1 = -(1+c)*vc+d_a/2*sqrt(1+c^2)
  arg2 = -(1+c)*vc+d_a/2/c*sqrt(1+c^2)
  
  rval = pnorm (arg1) + pnorm (arg2) - Heaviside (c)
  
  return (rval)
}

# TontData.R
c1 <- array(dim=c(5,2));d_a1 <- array(dim=c(5,2))

c1[1,1] <- -0.132280361
d_a1[1,1] <- 1.197239295
c1[2,1] <- -0.086965135	
d_a1[2,1] <- 1.771175637
c1[3,1] <- -0.144441852	
d_a1[3,1] <- 1.481934876
c1[4,1] <- 0.080460161	
d_a1[4,1] <- 1.513756914
c1[5,1] <- 0.222558765	
d_a1[5,1] <- 1.740157217

c1[1,2] <- -0.081742476	
d_a1[1,2] <- 0.628125133
c1[2,2] <- 0.049764485	
d_a1[2,2] <- 0.973878556
c1[3,2] <- -0.132612623	
d_a1[3,2] <- 1.155870661
c1[4,2] <- 0.118222633	
d_a1[4,2] <- 1.620175716
c1[5,2] <- 0.078103299	
d_a1[5,2] <- 0.89288159

d_a1 <- t(d_a1)
c1 <- t(c1)

trapz = function(x, y) 
{ ### computes the integral of y with respect to x using trapezoidal integration. 
  idx = 2:length(x)
  return (as.double( (x[idx] - x[idx-1]) %*% (y[idx] + y[idx-1])) / 2)
}
```


## Main code and output
  
```{r}
npts <-  10000
for (i in 1:2) {
  for (j in 1:5) {
    C  <-  c1[i,j]
    da  <-  d_a1[i,j]
    ret <- GetLimits(da,C)
    LL <- ret$LL;UL <- ret$UL
    vc  <-  seq (LL, UL, length.out = npts)
    TPF  <-  TruePositiveFraction (vc, da, C)
    FPF <- FalsePositiveFraction (vc, da, C)
    FPF <- rev(FPF);TPF <- rev(TPF)
    df2 <- data.frame(FPF = FPF, TPF = TPF)
    # do integral numerically
    numAuc <- trapz(FPF, TPF)
    # Implement Eqn. 36 from Metz-Pan paper 
    rho <- -(1-C^2)/(1+C^2);sigma <- rbind(c(1, rho), c(rho, 1))
    lower <- rep(-Inf,2);upper <- c(-da/sqrt(2),0)
    aucProproc <- pnorm(da/sqrt(2)) + 2 * pmvnorm(lower, upper, sigma = sigma)
    aucProproc <-  as.numeric(aucProproc)
    cat("i = ", i,"j = ", j,"C = ", C, ", da = ", da, "aucProproc =", aucProproc, "Norm. Diff. = ", (aucProproc-numAuc)/aucProproc,"\n")
  }
}
```             


## Discussion
Note the close correspondence between the formula, Eqn. 36 in the Metz-Pan paper and the numerical estimate. As a historical note, Eqn. 31 and Eqn. 36 (they differ only in parameterizations) in the referenced publication are provided without proof – it was probably obvious to Prof Metz or he wanted to leave it to us "mere mortals" to figure it out, as a final parting gesture of his legacy. The author once put a significant effort into proving it and even had a bright graduate student from the biostatistics department work on it to no avail. The author has observed that these equations always yield very close to the numerical estimates, to within numerical precisions, so the theorem is correct empirically, but he has been unable to prove it analytically. It is left as an exercise for a gifted reader to prove/disprove these equations.