---
title: "Chapter 20, Vignette 5: CBM plots"
author: "Dev P. Chakraborty, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
fig_caption: yes
fig.width: 4
vignette: >
  %\VignetteIndexEntry{CBM ROC plots}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
  library(RJafroc)
  library(ggplot2)
  library(mvtnorm)
```

## Helper functions
```{r echo=FALSE}
CbmRocY <- function (x, mu, alpha) {
  y <- (1-alpha)*(1-pnorm(qnorm(1-x))) + alpha*(1-pnorm(qnorm(1-x)-mu))
  return(y)
}
```


## Main code and output
  
```{r, fig.show='hold', echo=FALSE}
FPF <- seq(0.0, 1, 0.001)
alphaArr <- c(0.2, 0.8);muArr <- c(1,3)
myLabel <- c("A", "B", "C", "D")
myLabelIndx <- 1
for (i in 1:2)
  for (j in 1:2) 
  {
    {
      alpha <- alphaArr[i]
      mu <- muArr[j]
      TPF <- CbmRocY(FPF, mu, alpha)
      rocPlot <- data.frame(FPF = FPF, TPF = TPF)
      plotRoc <- ggplot(rocPlot, aes(x = FPF, y = TPF)) + 
        geom_line(data = rocPlot) + 
        scale_x_continuous(expand = c(0, 0)) + 
        scale_y_continuous(expand = c(0, 0)) +
      ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
      print(plotRoc)
      cat("Fig.", myLabel[myLabelIndx-1], ":", "mu = ", mu, ", alpha = ", alpha, "\n")
    }
  }
```             


## Comments
Plots A - D show ROC curves predicted by the CBM model; the corresponding values of the $mu$ and $alpha$ parameters are indicated above the plots. For small $mu$ and/or $alpha$  the curve approaches the chance diagonal, consistent with the notion that if the lesion is not visible, performance can be no better than chance level.

## pdf plots
```{r, fig.show='hold', echo=FALSE}
FPF <- seq(0.0, 1, 0.001)
alphaArr <- c(0.2, 0.8);muArr <- c(1,3)
myLabel <- c("E", "F", "G", "H")
myLabelIndx <- 1
for (i in 1:2)
  for (j in 1:2) 
  {
    {
      alpha <- alphaArr[i]
      mu <- muArr[j]
      if (i == 1) {
        z1 <- seq(-3, 3, by = 0.01)
        z2 <- seq(-3, mu + 3, by = 0.01)
      } else {
        z1 <- seq(-3, 3, by = 0.01)
        z2 <- seq(-3, mu + 3, by = 0.01)
      }
      Pdf1 <- dnorm(z1)
      Pdf2 <- (1 - alpha) * dnorm(z2) + alpha * dnorm(z2, mu)
      df <- data.frame(
        z = c(z1, z2), pdf = c(Pdf1, Pdf2), 
        truth = c(rep('non', length(Pdf1)), 
                  rep('dis', length(Pdf2)))
      )
      
      cbmPdfs <- ggplot(df, aes(x = z, y = pdf, color = truth)) +
        geom_line(data = df) +
        scale_colour_manual(values=c("black","darkgrey")) +
        scale_x_continuous(expand = c(0, 0)) +
        scale_y_continuous(expand = c(0, 0)) +
        #theme(legend.position = c(0.5, 0.2)) +
        theme(legend.position = "none") +
        ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1

      print(cbmPdfs)
      cat("Fig.", myLabel[myLabelIndx-1], ":", "mu = ", mu, ", alpha = ", alpha, "\n")
      next
    }
  }
```  

## Comments
The dark line is the diseased distribution. The grey line is the non-diseased distribution. The bimodal diseased distribution is clearly evident in plots F and H. 

## likelihood ratio plots
```{r, fig.show='hold', echo=FALSE}
FPF <- seq(0.0, 1, 0.001)
alphaArr <- c(0.2, 0.8);muArr <- c(1,3)
myLabel <- c("I", "J", "K", "L")
myLabelIndx <- 1
for (i in 1:2)
  for (j in 1:2) 
  {
    {
      alpha <- alphaArr[i]
      mu <- muArr[j]
      z <- seq(-3, 5 + mu, by = 0.01) # may need to adjust limits to view detail of slope plot near zero

      slope <- ((1-alpha)*dnorm(-z) + alpha*dnorm(mu-z))/dnorm(-z) # same as likelihood ratio
      slopePlot <- data.frame(z = z, slope = slope)
      plotSlope <- ggplot(slopePlot, aes(x = z, y = slope)) + 
        geom_line()  +
        ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
      print(plotSlope)
      cat("Fig.", myLabel[myLabelIndx-1], ":", "mu = ", mu, ", alpha = ", alpha, "\n")
      next
    }
  }
```  

## Comments
Close examination of the region near the flat part shows it does not plateau at zero; rather the minimum is at 1 - $alpha$, explaining the non-zero limiting slope of the predicted curve near (1, 1).