---
title: "Chapter 30, Vignette 2: Significance testing for the ROI paradigm"
author: "Dev P. Chakraborty, PhD"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
bibliography: myRefs.bib
fig_caption: yes
fig.width: 4
vignette: >
  %\VignetteIndexEntry{ROI Significance Testing}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
  library(RJafroc)
```
## Note to self (10/29/19) !!!DPC!!!
The FOM and DeLong method implementations need checking with a toy dataset.



## Introduction
* For an ROI dataset `StSignificanceTesting()` automatically defaults to `method = "ORH"`, `covEstMethod = "DeLong"` and `FOM = "ROI"`.  

* The covariance estimation method is based on the original DeLong method [@RN112], which is valid only for the trapezoidal AUC, i.e. ROC data, as extended by [@RN1233] to ROI data, see formula below.  

* The essential differences from conventional ROC analyses are in the definition of the ROI figure of merit, see below, and the procedure developed by [@RN1233] for estimating the covariance matrix. Once the covariances are known, `method = "ORH"` can be applied to perform significance testing, as described in [@RN1450] and [@RN2680, Chapter 10].

## The ROI figure of merit  
Let ${{X}_{kr}}$ denote the rating for the r^th^ **lesion-containing** ROI in the k^th^ case and let $n_{k}^{L}$ be the total number of lesion-containing ROIs in the k^th^ case. Similarly, let ${{Y}_{kr}}$ denote the rating for the r^th^ **lesion-free** ROI in the k^th^ case and $n_{k}^{N}$ denote the total number of lesion-free ROIs in the k^th^ case. Let ${{N}_{L}}$ denote the total number of lesion-containing ROIs in the image set and ${{N}_{N}}$ denote the total number of lesion-free ROIs. These are given by ${{N}_{L}}=\sum\nolimits_{k}{n_{k}^{L}}$ and ${{N}_{N}}=\sum\nolimits_{k}{n_{k}^{N}}$. The ROI figure of merit $\theta$\ is defined by: \[\theta =\frac{1}{{{N}_{L}}{{N}_{N}}}\sum\nolimits_{k}{\sum\nolimits_{{{k}'}}{\sum\limits_{r=1}^{n_{k}^{L}}{\sum\limits_{{r}'=1}^{n_{k'}^{N}}{\psi ({{X}_{kr}},{{Y}_{{k}'{r}'}})}}}}\]    

The kernel function $\Psi(X,Y)$ is defined by:

\[\psi (X,Y)=\left\{ \begin{align}
  & 1.0\ \ \ \text{if}\ Y<X \\ 
 & 0.5\ \ \ \text{if}\ Y=X \\ 
 & 0.0\ \ \ \text{if}\ Y>X \\ 
\end{align} \right.\]


The ROIs are *effectively regarded as mini-cases* and one calculates the FOM as the Wilcoxon statistic considering the mini-cases as actual cases. The correlations between the ratings of ROIs on the same case are accounted for in the analysis.


## Calculation of the ROI figure of merit.

```{r}
UtilFigureOfMerit(datasetROI, FOM = "ROI")
fom <- UtilFigureOfMerit(datasetROI, FOM = "ROI")
```             

* If the correct `FOM` is not supplied, it defaults to `FOM = ROI`.   
* This is a 2-treatment 5-reader dataset.  
* For treatment 1, reader 1 the figure of merit is `r fom[1,1]`.  
* For treatment 2, reader 5 the figure of merit is `r fom[2,5]`.  
* Etc.  

## Significance testing

When `dataset$dataType == "ROI"` the FOM defaults to "ROI" (meaning the above formula) and the covariance estimation method defaults to `covEstMethod = "DeLong"`.

```{r}
ret <- StSignificanceTesting(datasetROI, FOM = "Wilcoxon")
str(ret)
```             

* While `ret` is a list with many (22) members, their meanings should be clear from the notation. As an example:  

+    The variance components are given by:

```{r}
ret$varComp
```             

### RRRC analysis  

```{r}
ret$FTestStatsRRRC$fRRRC
ret$FTestStatsRRRC$ndfRRRC
ret$FTestStatsRRRC$ddfRRRC
ret$FTestStatsRRRC$pRRRC
```             

* The F-statistic is `r ret$fRRRC`, with `ndf = 1` and `ddf` = `r ret$ddfRRRC`, which yields a p-value of `r ret$pRRRC`.   

* The confidence interval for the reader averaged difference between the two treatments is given by:
    
```{r}
ret$ciDiffTrtRRRC
```             

* The FOM difference (treatment 1 minus 2) is `r ret$ciDiffTrtRRRC$Estimate`, which is significant, p-value = `r ret$ciDiffTrtRRRC$PrGTt`, F-statistic = `r ret$FTestStatsRRRC$fRRRC`, ddf = `r ret$FTestStatsRRRC$ddfRRRC`. The confidence interval is (`r ret$ciDiffTrtRRRC$CILower`, `r ret$ciDiffTrtRRRC$CIUpper`).

### FRRC analysis  

```{r}
ret$FTestStatsFRRC$fFRRC
ret$FTestStatsFRRC$ndfFRRC
ret$FTestStatsFRRC$ddfFRRC
ret$FTestStatsFRRC$pFRRC
```             

* The F-statistic is `r ret$FTestStatsFRRC$fFRRC`, with `ndf = 1` and `ddf` = `Inf`, which yields a p-value of `r ret$FTestStatsFRRC$pFRRC`. 

*  The confidence interval for the reader averaged difference between the two treatments is given by:
    
```{r}
ret$ciDiffTrtFRRC
```             

### RRFC analysis  

```{r}
ret$FTestStatsRRFC$fRRFC
ret$FTestStatsRRFC$ndfRRFC
ret$FTestStatsRRFC$ddfRRFC
ret$FTestStatsRRFC$pRRFC

```             

* The F-statistic is `r ret$FTestStatsRRFC$fRRFC`, with `ndf = 1` and `ddf` = `r ret$FTestStatsRRFC$ddfRRFC`, which yields a p-value of `r ret$FTestStatsRRFC$pRRFC`. 

* The confidence interval for the reader averaged difference between the two treatments is given by:
    
```{r}
ret$ciDiffTrtRRFC
```             

## Summary  
TBA

## References  
