---
title: "PROPROC ROCs and pdfs"
author: "Dev P. Chakraborty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
fig.width: 4
vignette: >
  %\VignetteIndexEntry{PROPROC ROCs}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---
  
```{r setup, include = FALSE}
  knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
  )
  library(RJafroc)
  library(ggplot2)
  library(mvtnorm)
```

## Helper functions
```{r}
Transform2ab <- function (d_a, c){
  b <- -(c+1)/(c-1)
  a <- (d_a/sqrt(2))*sqrt(1+b^2)
  return( list(
    a = a,
    b = b
  ) )
}

GetLimits <- function (d_a, c)
{
  if (c < 0.0) {
    LL <-  d_a/4/c*sqrt(1+c^2)
    UL <-  10.0
  }
  if (c == 0.0) 
  {
    LL  <-  -10.0
    UL <-  10.0
  }
  if (c > 0.0) 
  {
    LL <-  -10.0
    UL <-  d_a/4/c*sqrt(1+c^2)
  }
  return( list(
    LL = LL,
    UL = UL
  ) )
}

Heaviside <- function (c)
{
  if (c < 0.0) rval <-  0.0 else rval  <-  1.0
  return (rval)
}

FalsePositiveFraction <- function (vc, d_a, c)
{
  
  arg1 <-  -(1-c)*vc-d_a/2*sqrt(1+c^2)
  arg2 <-  -(1-c)*vc+d_a/2/c*sqrt(1+c^2)
  
  rval = pnorm (arg1) + pnorm (arg2) - Heaviside (c)
  
  return (rval)
}

TruePositiveFraction <- function (vc, d_a, c)
{
  arg1 = -(1+c)*vc+d_a/2*sqrt(1+c^2)
  arg2 = -(1+c)*vc+d_a/2/c*sqrt(1+c^2)
  
  rval = pnorm (arg1) + pnorm (arg2) - Heaviside (c)
  
  return (rval)
}

# rocY.R
# x is rocX, i.e., FPF
rocY <- function (x, a, b) {
  y <- pnorm(a + b*qnorm(x))
  return(y)
}
```
  
## TBA
  TBA 
  
## TBA
  TBA 
  
```{r}
c1Arr <-   c(-0.1322804, 0.2225588); daArr  <-  c(1.197239, 1.740157)
myLabel <- c("A", "B", "C", "D")
myLabelIndx <- 1
for (i in 1:2)
{
  c1 <- c1Arr[i]
  da <- daArr[i]
  ret <- Transform2ab(da, c1)
  a <- ret$a;b <- ret$b
  if (i == 1) z <- seq(-3, 0, by = 0.01) # may need to adjust limits to view detail of slope plot
  if (i == 2) z <- seq(-3, 5, by = 0.01) # may need to adjust limits to view detail of slope plot
  
  FPF <- seq(0.0, 1, 0.001)
  TPF <- rocY(FPF, a, b)
  
  rocPlot <- data.frame(FPF = FPF, TPF = TPF)
  plotRoc <- ggplot(rocPlot, aes(x = FPF, y = TPF)) + 
    geom_line()  + 
   scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0))   +
    ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1

  slope <-b*dnorm(a-b*z)/dnorm(-z) # same as likelihood ratio
  
  slopePlot <- data.frame(z = z, slope = slope)
  p <- ggplot(slopePlot, aes(x = z, y = slope)) + 
    geom_line()  + 
    scale_x_continuous(expand = c(0, 0)) + 
    scale_y_continuous(expand = c(0, 0))  +
    ggtitle(myLabel[myLabelIndx]);myLabelIndx <- myLabelIndx + 1
  print(plotRoc);print(p)
}
```             

Plot A is for `c` =  -0.1322804, `da` = 1.197239 while plot C is for `c` = 0.2225588, `da` = 1.740157. Plots B and D are the corresponding slope plots as functions of the binormal model z-sample. In plot A, the slope is infinite near the origin and the curve approaches the upper-right corner with finite slope. The situation is reversed in plot C where the slope is finite near the origin and the curve approaches the upper-right corner with zero slope. 
