---
title: "RJafroc Documentation"
author: "Dev P. Chakraborty"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Quick Start: DBMH Analysis Excel Output}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(RJafroc)
```

## Introduction
This vignette illustrates signficance testing using the DBMH method. But, 
instead of the unwieldy output in *QuickStartDBMH.html*, it generates an 
Excel output file containing the following worksheets:

* Summary
* FOMs
* RRRC
* FRRC
* RRFC
* ANOVA


## Generating the Excel output file

This illustrates the `UtilOutputReport()` function. The significance testing method is specified as `"DBMH"`
and the figure of merit `FOM` is specified as "Wilcoxon". Note the option `ReportFileFormat = "xlsx"` telling
the function to create an Excel output file

```{r}
UtilOutputReport(dataset03, stMethod = "DBMH", FOM = "Wilcoxon", overwrite = TRUE, ReportFileFormat = "xlsx")
```             



